Scikit-learn: API design for machine learning

### 2013-06-19 paper.tex
minor trimming and copyedit

In \sklearn, model selection is supported in two distinct meta-estimators,
\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as inputs an
estimator (basic or composite), whose hyper-parameters must be optimized, and a
hyper-parameter grid, which defines the values to search. Optionally, they can
also take a cross-validation scheme and a score function.  The library provides
various such cross-validation schemes, including K-Fold (default), Leave-One-Out
and Bootstrap. The score function used by default is the \texttt{score} method
provided by the estimator. However, the library provides a large variety of
score functions that the user can choose instead,
including accuracy, AUC, $F_1$-score for classification;
$R^2$ score or mean squared error for regression
and the Rand index for clustering.

@@

To facilitate the adoption and the distribution of \sklearn,
the set of external dependencies is kept to a bare minimum:
only Python, NumPy and SciPy are required for a functioning installation.
Binary distributions of these are available for the major platforms.
Visualisation functionality depends on Matplotlib \citep{hunter2007matplotlib}
and/or Graphviz \citep{gansner2000},
but neither is required to do machine learning.
When feasible, external libraries are integrated into the codebase.
In particular, \sklearn includes modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!
and wrapped using Cython modules.

@@

In comparison to specialized languages for numeric and statistical programming
such as \textsc{matlab} and R \citep{r},
the use of Python as a host language has the distinct advantage
that it is a \textit{general purpose} language:
it has strong language and standard library support for such tasks as
string handling/text processing, interprocess communication, networking
and many of the other tasks that many machine learning programs
(whether academic or commercial)
will have to perform ``on the side''.
In addition, third-party libraries are available for virtually any task
that modern computer programs routinely perform.

Another package that deserves mention is the Gensim topic modeling toolkit
\citep{rehurek2010gensim}.
Although its stated goal is not machine learning,
it does exemplify a style of API design for scalable processing of ``big data''
that takes a different approach than the comparable models in \sklearn.
Gensim's way of dealing with large datasets is to use algorithms
that have $O(1)$ space complexity and can be updated online,
called \textit{document streaming} by \citeauthor{rehurek2010gensim}.
The API is designed around the Python concept of an \textit{iterable}
(supported in the language by a restricted form of co-routines called
\textit{generators}).
While the text vectorization part of \sklearn
also uses iterables to some extent,
the emphasis is still on (mini-)batch processing,
even in the truly stateless, O(1) memory vectorizers
that implement the hashing trick of \citet{weinberger2009}.
This way of processing, as argued earlier in Section~\ref{sec:arrays},
reduces various forms of overhead
and makes effective use of the abundance of memory in modern computers.
No attempt is made to hide this batch-oriented processing from the user,
allowing control over the amount of memory actually dedicated
to \sklearn algorithms.

>>>>

In \sklearn, model selection is supported in two distinct meta-estimators,
\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as inputs an
estimator (basic or composite), whose hyper-parameters must be optimized, and a
hyper-parameter grid, which defines the values to search. Optionally, they can
also take a cross-validation scheme and a score function.  The library provides
various such cross-validation schemes, including K-Fold (default), Leave-One-Out
and Bootstrap. The score function used by default is the \texttt{score} method
provided by the estimator. However, the library provides a large variety of
score functions that the user can choose from,
including accuracy and $F_1$-score for classification;
$R^2$ score or mean squared error for regression;
and the Rand index for clustering.

@@

To facilitate the installation and thus adoption of \sklearn,
the set of external dependencies is kept to a bare minimum:
only Python, NumPy and SciPy are required for a functioning installation.
Binary distributions of these are available for the major platforms.
Visualisation functionality depends on Matplotlib \citep{hunter2007matplotlib}
and/or Graphviz \citep{gansner2000},
but neither is required to do machine learning.
When feasible, external libraries are integrated into the codebase.
In particular, \sklearn includes modified versions of \textsf{LIBSVM} and \textsf{LIBLINEAR}
\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++!
and wrapped using Cython modules.

@@

In comparison to specialized languages for numeric and statistical programming
such as \textsc{matlab} and R \citep{r},
the use of Python as a host language has the distinct advantage
that it is a \textit{general purpose} language:
it has strong language and standard library support for such tasks as
string handling/text processing, interprocess communication, networking
and many of the other tasks that many machine learning programs
(whether academic or commercial)
will have to perform ``on the side''.
In addition, third-party libraries are available for many other routine tasks.

Another package that deserves mention is the Gensim topic modeling toolkit
\citep{rehurek2010gensim}.
Although its stated goal is not machine learning,
it does exemplify API design for scalable processing of ``big data''
that takes a different approach than comparable models in \sklearn.
Gensim's way of dealing with large datasets is to use algorithms
that have $O(1)$ space complexity and can be updated online,
called \textit{document streaming} by \citeauthor{rehurek2010gensim}.
The API is designed around the Python concept of an \textit{iterable}
(supported in the language by a restricted form of co-routines called
\textit{generators}).
While the text vectorization part of \sklearn
also uses iterables to some extent,
the emphasis is still on (mini-)batch processing,
even in the stateless, O(1) memory vectorizers
that implement the hashing trick of \citet{weinberger2009}.
This way of processing, as argued earlier in Section~\ref{sec:arrays},
reduces various forms of overhead
and makes effective use of the abundance of memory in modern computers.
We make no attempt to hide this batch-oriented processing from the user,
allowing control over the amount of memory actually dedicated
to \sklearn algorithms.
